---
title: "Hospital Spending"
author: 'Nikesh,Pooja, Runhua '
date: "November 7, 2017"
output:
  html_document: default
  pdf_document: default
---

##Instructed by Andrea Wiggins

## R Script - 8086-002 - Data to Decisions

## The research questions are as follows:
1. Finding out on which claim type did all the hospital across USA has spent more and also for which period?  
2. Finding the amount spent in each state and grouping it under highest and lowest claim states?  
3. Finding the leading hospital in terms of spending based on the US zones and also finding the differences in spending by hospital on zone level?  

## The content of this Assignment
1. Set the path of the data and load the file
2. Loading the library as needed
3. Subset and transform the data
4. Produce the summary table for the subsets 
5. Process of plot using ggplot2 packages
6. Interpret and explain the figure.


##1.1 Plot for which claim type did all the hospital across USA has spent more 

###Setting the path to the folder of the date set, loading the data and checking the data
```{r set path and load the data}
setwd("D:/Graduate/3rd Sem(Fall)/8086-002 From Data To Decision/Team Project/R-Script")
HospitalSpending <- read.csv ("Medicare_Hospital_Spending_by_Claim_Cleaned.csv", stringsAsFactors = FALSE)  
View(HospitalSpending) # view the data
```

###Loading the library
```{r packages}
library(doBy)
library(DT)
library(xtable)
library(ggplot2)
library(reshape2)
library(GGally)
windowsFonts(Times=windowsFont("TT Times New Roman"))
```

### Subset and transform the data
```{r subset and transform the data based on claim type}
# select three columns of Claim_type, ASPE_Hospital, POS_Hospital for the folowing analyses
claimtype <- HospitalSpending[, c("Claim_type" , "ASPE_Hospital", "POS_Hospital")]

#change claim_type to full name for analyses
claimtype$Claim_type <- factor(claimtype$Claim_type, levels=c("Carrier","DME", "HHA", "Hospice", "Inpatient", "Outpatient", "SNF","Total"), labels=c("Carrier", "Durable Medical Equipment","Home Health Agency","Hospice","Inpatient","Outpatient", "Skilled Nursing Facility", "Total"))
```

### Produce the summary table for the subsets

```{r  summary table for the subsets based on claim type}
# summary table for the hospital spending nationally by claim types ($).
claimTable <- do.call(cbind.data.frame, aggregate(ASPE_Hospital ~ Claim_type, data=claimtype, FUN = function(x) {
  c("Number"=format(round(length(x), 0), nsmall = 0), M=format(round(mean(x), 2), nsmall = 2), min=min(x),max=max(x), 
    se=format(round(sd(x)/sqrt(length(x)), 2), nsmall = 2),format(round(quantile(x,c(0.05, 0.25, 0.50, 0.75, 0.95)), 2), nsmall = 2), iqr=IQR(x)) })); names(claimTable) <- c("Claim Type", "Number", "Mean", "Minimum", "Maximum", "Se", "5% Quantile","25% Quantile","Median","75% Quantile", "95% Quantile", "IQR")
View(claimTable) 

claimTablenew <- claimTable[1:7,c(1,3, 6)]
colnames(claimTablenew) <- c("claimType", "Mean", "SE")
str(claimTablenew)
# change the factor to numeric
as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}
claimTablenew$Mean <- as.numeric.factor(claimTablenew$Mean)
claimTablenew$SE <- as.numeric.factor(claimTablenew$SE)
str(claimTablenew)
```

### Produce the bar chart with ggplot2
```{r  produce bar char with ggplot2 based on claim type}
ggplot(claimTablenew, aes(x=claimType, y=Mean, fill=claimType)) +
  geom_bar(position=position_dodge(.9), colour="black", stat="identity") +
  geom_errorbar(position=position_dodge(.9), color="red", width=.25, aes(ymin=claimTablenew$Mean - claimTablenew$SE*2, ymax=claimTablenew$Mean + claimTablenew$SE*2)) +
  scale_fill_manual(values=c("lightblue","green","orange","blue","pink","lightseagreen","mediumvioletred")) +
  labs(x = "Claim Type", y = "Average Hospital Spending Across USA ($)") +
  ggtitle("Average Hospital Spending for Each Claim Type\n") + 
  theme_bw()+
  theme(panel.grid.major = element_line(color="gray50", size=0.2),
        panel.grid.major.x=element_blank(), 
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank(),
        legend.position=c("bottom"),
        legend.direction = "horizontal", 
        legend.justification = 0.5, 
        legend.key = element_rect(fill="black"), legend.background = element_rect(fill="gray"),
        legend.text =element_text(family="Times", size=10, face="bold", color="gray20"),
        axis.title.x=element_text(family="Times", color="black",face="bold", size=10, vjust=-1),
        axis.title.y=element_text(family="Times", color="black", face="bold", size=10, hjust=0.5),
        axis.text.y = element_text(family="Times", size =10, face="bold", margin = unit(c(t = 0, r = 2.5, b = 0, l = 0), "mm")),
        axis.line.x = element_line(color = "gray", size = 0.5),
        axis.line.y = element_line(color = "black", size = 0.5)) + 
  guides(fill=guide_legend(nrow=2, byrow=TRUE))+
  scale_y_continuous(breaks=seq(0,20000, 500)) 
```

**Figure 1.** Average Hospital Spending for Each Claim Type Across USA in 2015.  

### Figure 1. Interpretation and explanation  
 Figure 1 represents the average hospital spending for each claim type such as carrier, Durable medical equipment, home health agency, hospice, impatient, outpatient and skilled nursing facility based on the hospital spending data in 2015 from the following [links](https://catalog.data.gov/dataset/medicare-hospital-spending-by-claim-61b57). 
 From Figure 1, we found that the average hospital spending on inpatient is the most, skilled nursing facility, then outpatient, the least is durable medical equipment, during Index Hospital Admission is the most, then 1 through 30 days After Discharge.  This pattern may be explained by multiple costs from doctor, nurse, equipment, room and usually longer time in hospital comparing to other claim types.   



##1.2 Finding out on which period did all the hospital across USA has spent more 

### Subset the data

```{r subset and transform the data based on period}
# select three columns of Period, ASPE_Hospital, POS_Hospital for the folowing analyses
head(HospitalSpending)
periodIHA <- HospitalSpending[, c("Period" , "ASPE_Hospital", "POS_Hospital")]
```


# Produce the summary table for the subsets

```{r  summary table for the subsets based on period}
# summary table for the hospital spending nationally by periods ($).
periodTable <- do.call(cbind.data.frame, aggregate(ASPE_Hospital ~ Period, data=periodIHA, FUN = function(x) {
  c("Number"=format(round(length(x), 0), nsmall = 0), M=format(round(mean(x), 2), nsmall = 2), min=min(x),max=max(x),     se=format(round(sd(x)/sqrt(length(x)), 2), nsmall = 2),format(round(quantile(x,c(0.05, 0.25, 0.50, 0.75, 0.95)), 2), nsmall = 2), iqr=IQR(x)) })); names(periodTable) <- c("Periods", "Number", "Mean", "Minimum", "Maximum", "SE", "5% Quantile","25% Quantile","Median","75% Quantile", "95% Quantile", "IQR")
View(periodTable)  
periodTablenew <- periodTable[c(1,2,4), c(1,3, 6)]
colnames(periodTablenew) <- c("Periods", "Mean", "SE")
str(periodTablenew)
# change the factor to numeric
as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}
periodTablenew$Mean <- as.numeric.factor(periodTablenew$Mean)
periodTablenew$SE <- as.numeric.factor(periodTablenew$SE)
str(periodTablenew)
```

### Produce the bar chart with error bar by ggplot2
```{r  produce bar char with ggplot2 based on period}
ggplot(periodTablenew, aes(x=Periods, y=Mean, fill=Periods)) +
  geom_bar(position=position_dodge(.9), colour="black", stat="identity", width =0.5) +
  geom_errorbar(position=position_dodge(0.5), color="red", width=.25, aes(ymin=periodTablenew$Mean - periodTablenew$SE*2, ymax=periodTablenew$Mean + periodTablenew$SE*2)) +
  scale_fill_manual(values=c("green","blue","mediumvioletred")) +
  labs(x = "Periods", y = "Average Hospital Spending Across USA ($)") +
  ggtitle("Average Hospital Spending for each period\n") + 
  theme_bw()+
  theme(panel.grid.major = element_line(color="gray50", size=0.2),
        panel.grid.major.x=element_blank(), 
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank(),
        legend.position=c("bottom"),
        legend.direction = "horizontal", 
        legend.justification = 0.5, 
        legend.key = element_rect(fill="black"), legend.background = element_rect(fill="gray"),
        legend.text =element_text(family="Times", size=10, face="bold", color="gray20"),
        axis.title.x=element_text(family="Times", color="black",face="bold", size=10, vjust=-1),
        axis.title.y=element_text(family="Times", color="black", face="bold", size=10, hjust=0.5),
        axis.text.y = element_text(family="Times", size =10, face="bold", margin = unit(c(t = 0, r = 2.5, b = 0, l = 0), "mm")),
        axis.line.x = element_line(color = "gray", size = 0.5),
        axis.line.y = element_line(color = "black", size = 0.5)) + 
  scale_y_continuous(breaks=seq(0,20000, 500)) 
```

**Figure 2.** Average Hospital Spending for Each Period Across USA in 2015. Notes: AfterIHA stands for 1 to 3 days Prior to Index Hospital Admission; BeforeIHA stands for 1 through 30 days After Discharge from Index Hospital Admission; During IHA stands for During Index Hospital Admission.  

### Figure 2 Interpretation and explanation  
Figure 2 represents the average hospital spending for each period such as 1 to 3 days Prior to Index Hospital Admission; 1 through 30 days After Discharge from Index Hospital Admission and During Index Hospital Admission based on the hospital spending data from the following [links](https://catalog.data.gov/dataset/medicare-hospital-spending-by-claim-61b57). 
From Figure 2, we found that the average hospital spending on during Index Hospital Admission is the most, then 1 through 30 days After Discharge 
 from Index Hospital Admission, the least one is1 to 3 days Prior to Index Hospital Admission. This pattern may be due to multiple costs from doctor, nurse, equipment, room in hospital comparing to other periods, which are considerable expensive. 
  
  
 
 
 
 
##1.3 Finding the leading hospital in terms of spending based on the US zones and also finding the differences in spending by hospital on zone level?    


```{r include=FALSE, warning=FALSE}
setwd("D:/Graduate/3rd Sem(Fall)/8086-002 From Data To Decision/Team Project/R-Script")
hosp <- read.csv("Medicare_Hospital_Spending_by_Claim_Cleaned.csv")

#Removing "NA" values
hosp[is.na(hosp)] <- 0

# Creating the subset by filtering the values based on the column "Period"
hosp <- subset(hosp, Period == "ComEpisode")

# Grouping the state into Region
# North East Region
hosp_northeast <- subset(hosp, State %in% c("CT", "ME", "MA", "NH", "RI", "VT", "NJ", "PA", "NY"))
View(hosp_northeast)

#Midwest Region
hosp_midwest <- subset(hosp, State %in% c("IN", "IL", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD"))

#South Region
hosp_south <- subset(hosp, State %in% c("DE", "DC", "FL", "GA", "MD", "NC", "SC", "VA", "WV", "AL", "KY", "MS", "TN", "AR", "LA", "OK", "TX"))


#West Region
hosp_west <- subset(hosp, State %in% c("AZ", "CO", "ID", "NM", "MT", "UT", "NV", "WY", "AK", "CA", "HI", "OR", "WA"))
```
```{r include=FALSE,warning=FALSE}
library(scales)
fancy_scientific <- function(l) {
     # turn in to character string in scientific notation
     l <- format(l, scientific = TRUE)
     # quote the part before the exponent to keep all the digits
     l <- gsub("^(.*)e", "'\\1'e", l)
     # turn the 'e+' into plotmath format
     l <- gsub("e", "%*%10^", l)
     # return this as an expression
     parse(text=l)
}
```
## Region wise Plotting
###North East

```{r echo=FALSE,warning=FALSE}

library(ggplot2)
ggplot(hosp_northeast, aes(x=State, y = ASPE_Hospital, color = State))+
  geom_histogram(stat = "Identity")+
  xlab("Hospital Name")+
  ylab("Average Spending Per Episode Hospital")+
  scale_y_continuous(labels=fancy_scientific)
  ggtitle("Graph for highest spending per hospital in North East Region")
```
###Mid West
```{r echo=FALSE,warning=FALSE}
ggplot(hosp_midwest, aes(x=State, y = ASPE_Hospital, color = State))+
  geom_histogram(stat = "Identity")+
  xlab("Hospital Name")+
  ylab("Average Spending Per Episode Hospital")+
  scale_y_continuous(labels=fancy_scientific)
  ggtitle("Graph for highest spending per hospital in Midwest Region")
```
###South
```{r echo=FALSE,warning=FALSE}
ggplot(hosp_south, aes(x=State, y = ASPE_Hospital, color = State))+
  geom_histogram(stat = "Identity")+
  xlab("Hospital Name")+
  ylab("Average Spending Per Episode Hospital")+
  scale_y_continuous(labels=fancy_scientific)
  ggtitle("Graph for highest spending per hospital in South Region")
```
###West
```{r echo=FALSE,warning=FALSE}
ggplot(hosp_west, aes(x=State, y = ASPE_Hospital, color = State))+
  geom_histogram(stat = "Identity")+
  xlab("Hospital Name")+
  ylab("Average Spending Per Episode Hospital")+
  scale_y_continuous(labels=fancy_scientific)
  ggtitle("Graph for highest spending per hospital in West Region")
```

```{r include=FALSE, warning=FALSE}
#Assigning the dataframes to new dataframes.
NE <- hosp_northeast
MW <- hosp_midwest
S <- hosp_south
W <- hosp_west

#Add an additonal col called Region for all the subset data frame
NE$Region <- "North East"
MW$Region <- "Mid West"
S$Region <- "South"
W$Region <- "West"


# Sorting the column ASPE_Hospital in the asending order
NE_sort <- NE[order(NE$ASPE_Hospital, decreasing = TRUE),]
head(NE_sort)

MW_sort <- MW[order(MW$ASPE_Hospital, decreasing = TRUE),]
head(MW_sort)

S_sort <- S[order(S$ASPE_Hospital, decreasing = TRUE),]
head(S_sort)

W_sort <- W[order(W$ASPE_Hospital, decreasing = TRUE),]
head(W_sort)

# Assigning the values for comparision
NE_Firstrow <- head(NE_sort,1)
MW_Firstrow <- head(MW_sort,1)
S_Firstrow <- head(S_sort,1)
W_Firstrow <- head(W_sort,1)

#Bind the rows from multiple dataframes into single data frame
Region <- rbind(NE_Firstrow,MW_Firstrow,S_Firstrow,W_Firstrow)
head(Region)
#Select only the required columns
#Region <- Region[,c(1,3,6,13)]
Region <- Region[,c("Hospital_Name","State","ASPE_Hospital","Region")]
#Combining the Column State and Region to a new Column.
Region$Zone <- paste(Region$State, "\n", Region$Region)

#Combinging Hospital Name and Zone.
Region$Hospital <- paste(Region$Hospital_Name, "\n", Region$Zone)
```

```{r include=FALSE}
library(ggplot2)
library(ggthemes)
library(scales)
p <- ggplot(Region, aes(x=Hospital, y= ASPE_Hospital,fill=Hospital, label=ASPE_Hospital))+
  geom_histogram(stat="Identity")
```


## 1.3.2 Highest spending hospital in USA
### Finding the leading hospital in terms of spending based on the US zones and also finding the differences in spending by hospital on zone level?

###Graph for answering the above question
```{r echo=FALSE}

p + theme_solarized(light = FALSE) +
  scale_colour_solarized("red")+
  theme(axis.text.x = element_blank())+
  xlab("Hospital Name and Zones")+
  ylab("Average Spending Per Episode Hospital")+
  ggtitle("Highest spending hospital in USA")+
  geom_text(size = 3, position = position_stack(vjust = 1.05),color="red")
```

```{r}
library(ggmap)
write.csv(Region,"Region.csv")
r <- read.csv("Region.csv")
geocodes <- geocode(as.character(r$State))
r1 <- data.frame(r[,1:7],geocodes)
head(r1)
```


##Interpretation and explanation
Details about all the hospitals in USA who spend on patients over various claim types per year is present in particular dataset. Based on regions, We wanted to find the hospitals who spent more money on patients.  

Graph below, represents highest spending hospital by region.  

From the graph we can infer that, **Foundation Surgical hospital from Texas** has spent more money on its patients in South region. Similarly **Hackensack University medical center**,**Kings daughters medical center** and **Los Angeles** hospitals have spent good amount on patients in their regions.  

If we compare between the highest spending hospital from the graph then **Foundation Surgical hospital from Texas** has spent more when compared to **Hackensack University medical center**.  

```{r include=FALSE}
all_states <- map_data("state")
s <- ggplot()
s <- s + geom_polygon( data=all_states, aes(x=long, y=lat, group = group),colour="white", fill="grey10" )
```
#Same result represented in different way using MAP

```{r echo=FALSE}
s1 <- s +  geom_jitter( data=r1, position=position_jitter(width=0.5, height=0.5), aes(x=lon, y=lat, size = ASPE_Hospital, color=Hospital_Name))+
  scale_color_discrete(name="Hospital")+
  geom_text( data=r1, hjust=0.5, vjust=-2, aes(x=lon, y=lat, label=Hospital_Name), colour="magenta2", size=2 )+
  scale_size_continuous(range = c(5,9),guide = FALSE)+
  theme(panel.background = element_rect(fill = 'white', colour = 'red'))+
  theme(legend.position="bottom")+
  theme(legend.text=element_text(size=5))
s1
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Region}
summary(Region)
```



2) Finding the amount spent in each state and grouping it under highest and lowest claim states? 

# Finding the amount spent in each state and grouping it under highest and lowest claim states 

## Setting the path to the folder of the date set and loading the data 

```{r}

setwd("C:/Users/hp/Desktop")
StateSpending <- read.csv("Medicare_Hospital_Spending_by_Claim_Cleaned.csv", stringsAsFactors = FALSE)  
head(StateSpending) # check the first six rows of the data
View(StateSpending) # view the data
nrow(StateSpending) # check how many rows
ncol(StateSpending) # check how many columns

```

## Removing "NA" values

```{r}
StateSpending[is.na(StateSpending)] <- 0
```

## Creating the subset by filtering the values based on the column "Period"

```{r}
StateSpending <- subset(StateSpending, Period == "ComEpisode")
```

## Grouping the state into Region
## North East Region
```{r}
StateSpending_northeast <- subset(StateSpending, State %in% c("CT", "ME", "MA", "NH", "RI", "VT", "NJ", "PA", "NY"))
```


## Midwest Region
```{r}
StateSpending_midwest <- subset(StateSpending, State %in% c("IN", "IL", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD"))
```

## South Region

```{r}
StateSpending_south <- subset(StateSpending, State %in% c("DE", "DC", "FL", "GA", "MD", "NC", "SC", "VA", "WV", "AL", "KY", "MS", "TN", "AR", "LA", "OK", "TX"))
```


## West Region
```{r}
StateSpending_west <- subset(StateSpending, State %in% c("AZ", "CO", "ID", "NM", "MT", "UT", "NV", "WY", "AK", "CA", "HI", "OR", "WA"))
```

## Creating a subset with only Hospital_Name, State, ASPE_State
```{r}

(StateSpending <- StateSpending[, c("Hospital_Name", "State", "ASPE_State")])

```

## Creating a subset with only Hospital_Name, State, ASPE_State Regionwise
```{r}

(StateSpending_northeast <- StateSpending_northeast[, c("Hospital_Name", "State", "ASPE_State")])

(StateSpending_midwest <- StateSpending_midwest[, c("Hospital_Name", "State", "ASPE_State")])

(StateSpending_south <- StateSpending_south[, c("Hospital_Name", "State", "ASPE_State")])

(StateSpending_west <- StateSpending_west[, c("Hospital_Name", "State", "ASPE_State")])

```

## Aggregating the total expenditure statewise

```{r}

AggStateSpending<- aggregate(StateSpending$ASPE_State~StateSpending$State, data = StateSpending, sum)

```
## Loading the Library
```{r}
library(ggplot2)

```

## Changing the Column names

```{r}

colnames(AggStateSpending) <- c("State", "ASPE_State")

colnames(StateSpending_northeast) <- c("State", "ASPE_State")

colnames(StateSpending_south) <- c("State", "ASPE_State")

colnames(StateSpending_midwest) <- c("State", "ASPE_State")

colnames(StateSpending_west) <- c("State", "ASPE_State")


```

## Subsetting the Max and Min of the Aggregate State Spendings

```{r}
MaxAggStateSpending <- 
AggStateSpending[max(State ~ ASPE_State),]


MinAggStateSpending <- 
AggStateSpending[min(State ~ ASPE_State),]

```

## Statewise Plotting of the Spending

```{r}

ggplot(AggStateSpending, aes(x=State, y = ASPE_State, color = State))+
  + geom_histogram(stat = "identity")+
  + xlab("State")+
  + ylab("Average Spending Per Episode State")+
  + ggtitle("Graph for spending per state")
  
  ```

## Function for converting exponential to numeric

```{r}
fancy_scientific <- function(l) {
  +     # turn in to character string in scientific notation
    +     l <- format(l, scientific = TRUE)
    +     # quote the part before the exponent to keep all the digits
      +     l <- gsub("^(.*)e", "'\\1'e", l)
      +     # turn the 'e+' into plotmath format
        +     l <- gsub("e", "%*%10^", l)
        +     # return this as an expression
          +     parse(text=l)
        + }
```

## Reploting the graph after changing the expontential values to numbers
```{r}
ggplot(AggStateSpending, aes(x=State, y = ASPE_State, color = State))+
  + geom_histogram(stat = "identity")+
  + xlab("State")+
  + ylab("Average Spending Per Episode State")+
  + scale_y_continuous(labels=fancy_scientific)+
  + ggtitle("Graph for highest spending per State")
```

# Ordered Bar graph
```{r}
ggplot(AggStateSpending, aes(x=State, y=ASPE_State)) + 
  geom_bar(stat="identity", width=.5, fill="tomato3") + 
  labs(title="Ordered Statewise Spending", 
       subtitle="State Vs ASPE_State", ) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6))
  
```


#Reploting the graph after changing the expontential to numbers

```{r}
ggplot(StateSpending_northeast, aes(x=State, y = ASPE_State, color = State))+
  + geom_histogram(stat = "identity")+
  + xlab("State")+
  + ylab("Average Spending Per Episode State")+
  + scale_y_continuous(labels=fancy_scientific)+
  + ggtitle("Graph for highest spending per State in North East Region")

ggplot(StateSpending_midwest, aes(x=State, y = ASPE_State, color = State))+
  geom_histogram(stat = "Identity")+
  xlab("State")+
  ylab("Average Spending Per Episode State")+
  scale_y_continuous(labels=fancy_scientific)
ggtitle("Graph for highest spending per State in Midwest Region")

ggplot(StateSpending_south, aes(x=State, y = ASPE_State, color = State))+
  geom_histogram(stat = "Identity")+
  xlab("State")+
  ylab("ASPE_State")+
  scale_y_continuous(labels=fancy_scientific)
ggtitle("Graph for highest spending per State in South Region")

ggplot(StateSpending_west, aes(x=State, y = ASPE_State, color = State))+
  geom_histogram(stat = "Identity")+
  xlab("State")+
  ylab("ASPE_State")+
  scale_y_continuous(labels=fancy_scientific)
ggtitle("Graph for highest spending per State in West Region")

```
